A Data-Driven Framework for Neural Field Modelling

### 2010-10-26 ltx/BrainIDEPoster/BrainIDE_Poster.tex
edits to the poster

      \vskip2ex

      \begin{block}{Neural Field Model}
        \rmfamily{The model's temporal dynamics are described by 
		\begin{equation}
			\label{SpikesToPotential} v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'}, 
		\end{equation}
		where the post-synaptic response kernel $h(t)$ is described by 
		\begin{equation}
			\label{SynapticRespKernel} h(t) = \eta(t)\exp{\left(-\zeta t\right)}, 
		\end{equation}
		and $\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function. Spatial interactions between cortical populations are described by 
		\begin{equation}
			\label{RateBasedInteractions} g\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)\textrm{d}\mathbf{r}'}, 
		\end{equation}
		where $f(\cdot)$ is a sigmoidal firing rate function, $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing a cortical sheet or surface. Substituting equation~\ref{RateBasedInteractions} into equation~\ref{SpikesToPotential} gives the spatiotemporal model 
		\begin{equation}
			\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) =
			\int_{-\infty}^t 
			h\left(t - t'\right) \int_\Omega
			w\left(\mathbf{r},\mathbf{r}'\right) 
			f\left( v\left( \mathbf{r}',t' \right)\right)
			\textrm{d}\mathbf{r}'dt'.
		\end{equation}		
The model can be discretised in time and written as an integro-difference equation in the form
% \begin{alertblock}{Neural Field IDE}
\begin{equation}
	\label{DiscreteTimeModel} 
	v_{t+1}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r}'\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\textrm{d}\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
% \end{alertblock}
where $T_s$ is the time step, $\xi = 1-T_s\zeta $ and $e_t(\mathbf{r})$ is an $i.i.d.$ disturbance such that $e_t(\mathbf{r})\sim\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$. Here $\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$ denotes a zero mean spatial Gaussian process with covariance function $\gamma(\mathbf{r}-\mathbf{r}')$.

The mapping between the membrane voltage and the electrophysiological data is
	\begin{equation}
	    \label{eq:ObservationEquation}
		\mathbf{y}_t =
		\int_{\Omega}{
		    m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)
		\textrm{d}\mathbf{r}'} + 
		\boldsymbol{\varepsilon}_t, 
	\end{equation}
	where $\mathbf{r}_n$ defines the electrode positions, $n=1,...,N$ indexes the electrodes and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\boldsymbol{\Sigma}_{\varepsilon}\right)$. $\mathcal{N}\left(0,\boldsymbol{\Sigma}_{\varepsilon}\right)$ denotes the multivariate normal distribution with mean zero and covariance matrix $\boldsymbol{\Sigma}_{\varepsilon}$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the sensor pick-up geometry and is defined by 
	\begin{equation}
		m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)},
	\end{equation}
	where $\sigma_m$ sets the sensor width.}
    \end{block}

@@

\begin{block}{State-Space Model}
	\begin{figure}	
	\begin{center}
	  \includegraphics[width=7in, scale = 10]{Figure2.eps}
	\end{center}
	\caption{{\bf Example of a one-dimensional field decomposition.} A continuous, one-dimensional field decomposed by a finite number of basis functions scaled by the state vector. The dashed line depicts the field, the black solid lines shows the Gaussian basis functions and the vertical lines show the position and amplitude of the states.} 
	\label{fig:Figure2}	
	\end{figure}
	\rmfamily{To implement standard estimation techniques, we use a set of Gaussian basis functions to decompose the field \cite{Dewar2009}. The field decomposition is described by 
	\begin{equation}
		\label{DefFieldDecomp} v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t, 
	\end{equation}
	where $\mathbf{x}_t$ is the state vector that scales the field basis functions $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$. The field basis functions are described by
	\begin{equation}\label{eq:FieldBasisFunction}
		\boldsymbol\phi\left(\mathbf{r}-\mathbf{r}'\right) =
	\exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_{\phi}^2}\right)}. 
	\end{equation}
	The connectivity kernel can also be decomposed as 
	\begin{equation}\label{DefKernelDecomp}
		 w\left(\mathbf{r},\mathbf{r}'\right) =\boldsymbol{\psi}^\top\left(\mathbf{r},\mathbf{r}'\right) \boldsymbol{\theta},
	\end{equation}
	where $\boldsymbol{\psi}(\mathbf{r},\mathbf{r}')$ is a vector of Gaussian basis functions and $\boldsymbol{\theta}$ is a vector of scaling parameters. Now defining the matrices
	\begin{align}
		\boldsymbol{\Gamma} &\triangleq \int_\Omega {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi} ^{\top}\left(\mathbf{r}\right)\textrm{d}\mathbf{r}} \label{eq:DefGamma} \\ 
		\boldsymbol{\Psi}(\mathbf{r}') &\triangleq T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top}\label{eq:DefPsi} (\mathbf{r}'-\mathbf{r})\textrm{d}\mathbf{r}},
	\end{align}
	where $\Gamma$ is a $L\times L$ matrix and $\Psi$ is $L \times n_{\theta}$ matrix, where $L$ is the number of basis functions (states) and $n_{\theta}$ is the number of connectivity kernel basis functions. The state disturbance is defined as
	\begin{equation}\label{eq:Wt} 
		\mathbf{e}_t \triangleq \boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi} ( \mathbf{r} )e_t( \mathbf{r} )\textrm{d}\mathbf{r}},
	\end{equation}
	which is a zero mean normally distributed white noise process with covariance
	\begin{equation}
		\boldsymbol\Sigma_e =\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 
	\end{equation}
	By rearranging Eq.~\ref{DiscreteTimeModel} and substituting in Eqs. \ref{DefFieldDecomp}, \ref{eq:FieldBasisFunction}, \ref{eq:DefGamma}, \ref{eq:DefPsi} and \ref{eq:Wt} we get the following state-space model
	% \begin{alertblock}{State-Space Model}
	\begin{align}
		\mathbf{x}_{t+1} &= \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \boldsymbol{\theta} + \xi\mathbf{x}_t + \mathbf{e}_t \\
		\mathbf{y}_t &= \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t.
		\end{align} 	
	% \end{alertblock}
	where the observation matrix is 
	\begin{equation}
			\mathbf{C} = \left[
			\begin{array}{{ccc}} 
				c_{1,1} & \dots & c_{1,L} \\
				\vdots & \ddots & \vdots \\
				c_{N,1} & \dots & c_{N,L} 
			\end{array}
			\right] 
		\end{equation}
		and 
		\begin{equation}
			c_{i,j} = \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\phi}_j(\mathbf{r}')\textrm{d}\mathbf{r}'. 
		\end{equation}
	}
\end{block}

@@

      \begin{block}{Frequency Analysis}
		\begin{figure}
		\begin{center}
		\includegraphics[width=7in, scale = 10]{Figure4.eps}
		\end{center}
		\caption{{\bf Spatial frequency analysis of the neural field}. (\textbf{A}) The average (over time) power in dB of the spatial frequency of the neural field. (\textbf{B}) The average (over time) power in dB of the spatial frequency of the reconstructed neural field from the basis function decomposition. The reconstructed field shows a narrower bandwidth as a result of the band limiting properties of the sensors and basis functions.}
		\label{fig:Figure4}
		\end{figure}
		\rmfamily{Given a band-limited field, the distance between adjacent sensors, $\Delta_y$, must satisfy 
		\begin{equation}
			\label{eq:MinimumSensorDistance} \Delta_y \leq \frac{1}{2\rho_y\boldsymbol{\nu}_{c}}, 
		\end{equation}
		where $\rho_y \in \mathbb{R} \ge 1$ is an oversampling parameter and $\boldsymbol{\nu}_c$ is a cutoff frequency (typically taken as the -3~dB point). This condition must be satisfied to avoid spatial spectral aliasing effects when reconstructing the hidden dynamic field, $v_t(\mathbf{r})$, using the sampled observations, $\mathbf{y}_t$.

@@

      \begin{block}{Results}
		\begin{figure}	
		\begin{center}
		  \includegraphics[width=7in, scale = 10]{Figure7.eps}
		\end{center}
		\caption{{\bf Histograms of the parameter estimates over 100
		realisations}. Solid lines show the actual values of the parameters and the dotted lines show the means of estimated parameters. (\textbf{A}) The synaptic time constant estimates. (\textbf{B}) The central excitatory connectivity basis function parameter estimates. (\textbf{C}) The surround inhibition connectivity basis function parameter estimates. (\textbf{D}) The longer range excitatory connectivity basis function parameter estimates.} 
		\label{fig:Figure7}	
		\end{figure}
		\begin{figure}
		\begin{center}
		\includegraphics[width=7in, scale = 10]{Figure8.eps}
		\end{center}
		\caption{{\bf Cross-section of connectivity kernel estimate}. The actual kernel is shown by the solid line, the mean kernel estimate (over 100 realisations) is shown by the dotted line and the 95~\% confidence interval is shown by the shaded grey region.}
		\label{fig:Figure8}
		\end{figure}

>>>>

      \vskip2ex
% \begin{block}{Notation}
% 	\rmfamily{
% 	
% 	}
% \end{block}
      \begin{block}{Neural Field Model}
        \rmfamily{The model's temporal synaptic dynamics are described by 
		\begin{align}
			\label{SpikesToPotential} v\left( {\mathbf{r},t} \right) &= \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'} \\ 
			h(t) &= \eta(t)\exp{\left(-\zeta t\right)}, 
		\end{align}
		where $g(\mathbf{r},t)$ is the input firing rate, $\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function. Interactions between cortical populations are described by 
		\begin{equation}
			\label{RateBasedInteractions} g\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)\textrm{d}\mathbf{r}'}, 
		\end{equation}
		where $f(\cdot)$ is a sigmoidal firing rate function, $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing a cortical sheet or surface. Substituting equation~\ref{RateBasedInteractions} into equation~\ref{SpikesToPotential} gives the spatiotemporal model 
		\begin{equation}
			\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) =
			\int_{-\infty}^t 
			h\left(t - t'\right) \int_\Omega
			w\left(\mathbf{r},\mathbf{r}'\right) 
			f\left( v\left( \mathbf{r}',t' \right)\right)
			\textrm{d}\mathbf{r}'dt'.
		\end{equation}
The discrete time integro-difference equation form of the model is
\begin{alertblock}{Neural Field IDE}
\begin{equation}
	\label{DiscreteTimeModel} 
	v_{t+1}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r}'\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\textrm{d}\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
\end{alertblock}
where $T_s$ is the time step, $\xi = 1-T_s\zeta $ and $e_t(\mathbf{r})\sim\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$.

The mapping between the neural field and the electrophysiological data is
	\begin{equation}
	    \label{eq:ObservationEquation}
		\mathbf{y}_t =
		\int_{\Omega}{
		    m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)
		\textrm{d}\mathbf{r}'} + 
		\boldsymbol{\varepsilon}_t, 
	\end{equation}
	where $\mathbf{r}_n$ defines the electrode positions, $n=1,...,N$ indexes the electrodes and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\boldsymbol{\Sigma}_{\varepsilon}\right)$. The output kernel, $m(\mathbf{r}-\mathbf{r}')$, governs the sensor pick-up geometry and is defined by 
	\begin{equation}
		m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)},
	\end{equation}
	where $\sigma_m$ sets the sensor width.}
    \end{block}

@@

\begin{block}{State-Space Model}
	\begin{figure}	
	\begin{center}
	  \includegraphics[width=8in, scale = 10]{Figure2.eps}
	\end{center}
	\caption{{\bf Example of a one-dimensional field decomposition.} The dashed line depicts the field, the black solid lines shows the Gaussian basis functions and the vertical lines show the position and amplitude of the states.} 
	\label{fig:Figure2}	
	\end{figure}
	\rmfamily{To implement standard estimation techniques, we use a set of Gaussian basis functions to decompose the field \cite{Dewar2009}. The field decomposition is described by 
	\begin{equation}
		\label{DefFieldDecomp} v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t, 
	\end{equation}
	where $\mathbf{x}_t$ is the state vector that scales the field basis functions $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$. 
	The connectivity kernel can also be decomposed as 
	\begin{equation}\label{DefKernelDecomp}
		 w\left(\mathbf{r},\mathbf{r}'\right) =\boldsymbol{\psi}^\top\left(\mathbf{r},\mathbf{r}'\right) \boldsymbol{\theta},
	\end{equation}
	where $\boldsymbol{\psi}(\mathbf{r},\mathbf{r}')$ is a vector of Gaussian basis functions and $\boldsymbol{\theta}$ is a vector of scaling parameters. Now defining the matrices
	\begin{align}
		\boldsymbol{\Gamma} &\triangleq \int_\Omega {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi} ^{\top}\left(\mathbf{r}\right)\textrm{d}\mathbf{r}} \label{eq:DefGamma} \\ 
		\boldsymbol{\Psi}(\mathbf{r}') &\triangleq T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top}\label{eq:DefPsi} (\mathbf{r}'-\mathbf{r})\textrm{d}\mathbf{r}},
	\end{align}
	where $\Gamma$ is a $L\times L$ matrix and $\Psi$ is $L \times n_{\theta}$ matrix, where $L$ is the number of basis functions (states) and $n_{\theta}$ is the number of connectivity kernel basis functions. The state disturbance is defined as
	\begin{equation}\label{eq:Wt} 
		\mathbf{e}_t \triangleq \boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi} ( \mathbf{r} )e_t( \mathbf{r} )\textrm{d}\mathbf{r}},
	\end{equation}
	which is a zero mean normally distributed white noise process with covariance
	\begin{equation}
		\boldsymbol\Sigma_e =\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 
	\end{equation}
	By rearranging Eq.~\ref{DiscreteTimeModel} and substituting in Eqs. \ref{DefFieldDecomp}, \ref{eq:FieldBasisFunction}, \ref{eq:DefGamma}, \ref{eq:DefPsi} and \ref{eq:Wt} we get the following state-space model
	\begin{alertblock}{State-Space Model}
	\begin{align}
		\mathbf{x}_{t+1} &= \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \boldsymbol{\theta} + \xi\mathbf{x}_t + \mathbf{e}_t \\
		\mathbf{y}_t &= \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t.
		\end{align} 	
	\end{alertblock}
	where the observation matrix is 
	\begin{equation}
			\mathbf{C} = \left[
			\begin{array}{{ccc}} 
				c_{1,1} & \dots & c_{1,L} \\
				\vdots & \ddots & \vdots \\
				c_{N,1} & \dots & c_{N,L} 
			\end{array}
			\right] 
		\end{equation}
		and 
		\begin{equation}
			c_{i,j} = \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\phi}_j(\mathbf{r}')\textrm{d}\mathbf{r}'. 
		\end{equation}
	}
\end{block}

@@

      \begin{block}{Frequency Analysis}
		\begin{figure}
		\begin{center}
		\includegraphics[width=8in, scale = 10]{Figure4.eps}
		\end{center}
		\caption{{\bf Spatial frequency analysis of the neural field}. (\textbf{A}) The average (over time) power in dB of the spatial frequency of the neural field. (\textbf{B}) The average (over time) power in dB of the spatial frequency of the reconstructed neural field from the basis function decomposition.}
		\label{fig:Figure4}
		\end{figure}
		\rmfamily{Given a band-limited field, the distance between adjacent sensors, $\Delta_y$, must satisfy 
		\begin{equation}
			\label{eq:MinimumSensorDistance} \Delta_y \leq \frac{1}{2\rho_y\boldsymbol{\nu}_{c}}, 
		\end{equation}
		where $\rho_y \in \mathbb{R} \ge 1$ is an oversampling parameter and $\boldsymbol{\nu}_c$ is a cutoff frequency (typically taken as the -3~dB point). This condition must be satisfied to avoid spatial spectral aliasing effects when reconstructing the hidden dynamic field, $v_t(\mathbf{r})$, using the sampled observations, $\mathbf{y}_t$.

@@

      \begin{block}{Results}
		\begin{figure}	
		\begin{center}
		  \includegraphics[width=8in, scale = 10]{Figure7.eps}
		\end{center}
		\caption{{\bf Histograms of the parameter estimates over 100
		realisations}. Solid lines show the actual parameters and the dotted lines show the means of estimated parameters. (\textbf{A}) The synaptic time constant. (\textbf{B}) The central excitatory connectivity parameter. (\textbf{C}) The surround inhibition connectivity parameter. (\textbf{D}) The longer range excitatory connectivity parameter.} 
		\label{fig:Figure7}	
		\end{figure}
		\begin{figure}
		\begin{center}
		\includegraphics[width=8in, scale = 10]{Figure8.eps}
		\end{center}
		\caption{{\bf Cross-section of connectivity kernel estimate}. The actual kernel is shown by the solid line, the mean kernel estimate (over 100 realisations) is shown by the dotted line and the 95~\% confidence interval is shown by the shaded grey region.}
		\label{fig:Figure8}
		\end{figure}
